---
title: "library() vs require()"
date: '2019-05-28'
output:
  word_document: default
  html_document:
    df_print: paged
header:
  caption: ''
  image: ''
editor_options:
  chunk_output_type: console
slug: library() vs require()
tags: []
categories: []
---


The use of `library()` or `require()` is interesting. Each have is own pros and cons. If we'll use the help for library (?library), we'll see that "library and require load and attach add-on packages". So, while practicality they do the same thing, they act differently. 

In both cases both functions load the packages with name package and put it on the search list. However, When we call `library()`, if the package doesn't exists a warning will be raised. In addition, the script will stop running.  
```{r, eval = FALSE}
library(somepackage)
```

```{r, echo=FALSE}
warning("Error in library(somepackage) : there is no package called ‘somepackage’)")

```

On the other hand, when we call `require()` the function will return `TRUE` if the package exists, and `FALSE` if the package doesn't exists. furthermore, unlike `library()`, the script will not stop running. 
```{r}
print(require(somepackage))
```

So, why do we care about these differences? 
Beside knowing what each function does, we can use them to our own advantage. 

It's considered better practice to use `library()` when we want to load packages outside of functions. This is because we would like to know if and where we have a bug as soon as possible. Calling a set of function which doesn't exists is certainly a bug. We'll use `require()` inside functions, as we can manipulate it's `TRUE`/`FALSE` output. For example, in the beginning of our script, we can write a function that will install all the necessary function if they aren't already installed [look here](https://amirdjv.netlify.com/post/r-and-rstudio-tricks-and-shortcuts/). We could also skip parts of a script if the packages aren't already installed, etc. 

Thank you for reading! 





