---
title: "Creating functions in R"
author: "Amir Djalovski"
date: '2019-01-08'
output: word_document
header:
  caption: ''
  image: ''
editor_options:
  chunk_output_type: console
slug: Creating functions in R
tags: []
categories: []
---

Building functions are one of the main building blocks in programming and in writing reproducible, easy to understand and readable code. In addition, it allows you to write what you need in a more efficient way and to detect errors more easily. Besides these important things, in most cases, your code will run faster. After I convinced you, I hope, we can start learning how to write functions.

Functions can do everything that you can code and ideally whenever you’re writing the same code more than twice you’ll make a function out of it.

To create a function we need to set the number of inputs the function we’ll take, what it will do, and to name the function (sometimes). Let’s start with the following example:
```{r, eval=TRUE}

square <- function(x) { 
  out <- x^2
  return(out)
  }

```

We assigned the function to `square`, thus whenever we'd like to call the function we'll use `square()`. Next, the function takes only one argument (`x`), the input will be squared and will get assigned to `out`. Lastly, using the `base R` function `return`, the function will return the value of `out`. 

Practically, we don't have to assign the function to `out` or use `return` in this example. However, getting used to using `return` is a good habit. Whenever we build a long function and we want to debug it, it's always nice to look for the `return` and see what the function suppose to return.

Now, we can test our function to see if it actually return the expected results
```{r}
square(2)
```

And we can also use our function over a vector
```{r}
myVec <- c(1:4)
print(myVec)

square(myVec)
```


We can even create more complicated function with two or three inputs (or more!). let's create a function which calculate the [Pythagorean theorem](https://en.wikipedia.org/wiki/Pythagorean_theorem). 
```{r}
pythagorean <- function(x, y) { 
  out <- x^2 + y^2
  return(out)
}
```

We have two inputs (`x` and `y`) and will add the square value of both of them. The result will be assigned to `out` and will be `reutrn`ed. 
We can test our function 
```{r}
pythagorean(2, 4) 
```


Thank you for reading! 




